---
title: "Inference Statistics Course Project 3"
author: "Mathias Barat"
date: "25/09/2020"
output:
  pdf_document: default
  html_document: default
---

# PART 1 - SIMULATION EXERCISE

## Overview

Investigate the exponential distribution in R and compare it with the Central limit theorem


## Simulation

```{r}
# Set-up my inputs:
lambda <- 0.2
mu <- 1/lambda
sd <- 1/lambda
n <- 40
n_simu <- 1000


# data simulation
set.seed(12345)
matrix_sample <- matrix(rexp(n*n_simu,lambda), n_simu, n)
mean_sample <- rowMeans(matrix_sample)


# summary
summary(mean_sample)
```
## Sample Mean


```{r}

mua <- mean(mean_sample)

deltamu <- mua - mu
deltamu
round (deltamu/mu*100,1)
```
The sample mean (`r mua`) is 0.6% lower than the theoretical mean (`r mu`). The difference is really minimal.

## Sample Variance

```{r}

Var_theo <- sd^2 / n
Var_sample <- var(mean_sample)

Var_theo
Var_sample

deltavar = Var_sample - Var_theo
var_perc <- round(deltavar/Var_theo*100,2)
```

The Theoretical Variance (`r Var_theo`) and is `r var_perc`% lower than the Sample Variance is (`r Var_sample`).

## Distribution

```{r}
library(ggplot2)

sd <- sqrt(Var_theo)

ggplot(as.data.frame(mean_sample), aes(as.data.frame(mean_sample)[,1]))+
        geom_histogram(aes(y=..density..), position="identity", fill="yellow", col="black")+
        geom_density(colour="red", size=3)+
        stat_function(fun = dnorm, colour = "green", args = list(mean = mu, sd = sd), size=3)+
        ggtitle ("Sample Means Distribution EXP")+
        labs(subtitle = "Fitting normal curve - Rastafari Style")+
        xlab("Sample mean")+
        ylab("Density")

```
As highlighted by the plot, we can see a small discrepancy between the sample distribution and the theoretical one.

## Conclusion

We can conclude that the simulation performed with the rexp R function has produced a random dataset globally normal.
