---
title: "Inference Statistics Course Project 3"
author: "Mathias Barat"
date: "25/09/2020"
output:
  pdf_document: default
  html_document: default
---

# PART 2

## Overview

Analyse the ToothGrowth dataset :

The tooth growth data set is the length of the odontoblasts (teeth) in each of 10 guinea pigs at different Vitamin C dosage levels with two delivery methods.

The procedure will consists in:
  - Doing Exploratory Data Analyses
  - Provide a summary
  - Perform confidence interval
  - State some conclusions.

## Load the Dataset

```{r}

library(datasets)
library(ggplot2)
mydata <- ToothGrowth

```

## EDA

### Structure of the dataset

```{r}
str(mydata)
```
The dataframe has 60 observations and 3 columns.

  - len : Tooth length
  - supp : Supplement Type 
      -> "OJ" for orange Juice
      -> "VC" for Ascorbic Acid (???)
  - dose : 3 levels of Vitamin C dosage (0.5, 1, 2 mg)

```{r}
sum(is.na(mydata))
```
There is no NA in the dataset.
```{r}
summary(mydata$len)
```
### Plot

A first blind graph to show roughly to have an idea of the content.

```{r}
qplot(data = mydata, y = len , facets = . ~ dose, )+
        aes(color = supp, fill= supp) +
        ggtitle ("Growth Tooth")+
        labs(subtitle = "Exploratory Data analysis")+
        xlab("Tooth Length")+
        ylab("Length")
```
We can directly observed that when the doses of Vitamin C are increasing the teeth are increasing also.

Let's summarize closely the dataset:

### Summarize the Tooth Length by Dose and Supp

```{r}
library(dplyr)

sum_tot <- mydata %>% 
    group_by(supp,dose) %>%
    summarize(mean_len_tooth=mean(len), sd_len_tooth=sd(len), count = n())
print(sum_tot)

```

### Summarize the Tooth Length by Supp only

```{r}
library(dplyr)

sum_supp <- mydata %>% 
    group_by(supp) %>%
    summarize(mean_len_tooth=mean(len), sd_len_tooth=sd(len), count = n())
print(sum_supp)

```

### Summarize by Dosage level

```{r}
sum_dose <- mydata %>% 
    group_by(dose) %>%
    summarize(mean_len_tooth=mean(len), sd_len_tooth=sd(len), count = n())

print(sum_dose)

```

Clearly, the teeth length means are greater when the doses of vitamin C increase.
Same observation when we administrate the treatment with the Orange Juice.

## Confidence Interval/Hypothesis

Let's run t.test for the different configurations possible of the data:

### Supplement Method Comparison

##### At all dosage levels:
```{r}

t.test(len ~ supp, paired=FALSE, var.equal=FALSE, data=mydata)

```
##### At 0.5mg dosage level:
```{r}

t.test(len ~ supp, paired=FALSE, var.equal=FALSE, data=mydata[mydata$dose==0.5,])

```
##### At 1mg dosage level:
```{r}

t.test(len ~ supp, paired=FALSE, var.equal=FALSE, data=mydata[mydata$dose==1,])

```
##### At 2mg dosage level:
```{r}

t.test(len ~ supp, paired=FALSE, var.equal=FALSE, data=mydata[mydata$dose==2,])

```

We can find a significant difference between the 2 supplement methods for the 0.5 and 1mg dosage. No significant difference at 2mg.

### Dosage Comparison

So we will compare the different dosage with OJ:

##### Compare 0.5 to 1 for OJ:
```{r}

t.test(len ~ dose, paired=FALSE, var.equal=FALSE, data=mydata[mydata$dose<2 & mydata$supp=="OJ",])

```
##### Compare 1 to 2 for OJ:
```{r}

t.test(len ~ dose, paired=FALSE, var.equal=FALSE, data=mydata[mydata$dose>0.5 & mydata$supp=="OJ",])


```
##### Compare 0.5 to 1 for VC:
```{r}

t.test(len ~ dose, paired=FALSE, var.equal=FALSE, data=mydata[mydata$dose<2 & mydata$supp=="VC",])

```
##### Compare 1 to 2 for VC:
```{r}

t.test(len ~ dose, paired=FALSE, var.equal=FALSE, data=mydata[mydata$dose>0.5 & mydata$supp=="OJ",])


```
For all the tests comparing the dosage, the confidence interval is always excluding 0. The differences between the dosage levels are significant.

## Statements of the study

### Conclusions

  - The Vitamin C is correlated to the tooth growth with high confidence (95%) and this whatever the supplement method.
  - The Orange Juice is providing better tooth growth at low dosage (=<1mg) than Ascorbic Acid. There is no significant difference at 2mg.
  
### Assumptions:
  - The measurement are not paired
  - The variances are not equal
  - The test subjects were randomly selected and independants.
